---
output: html_document
editor_options: 
  chunk_output_type: console
---
--
title: "R Notebook"
output: html_notebook
chunk_output_type: console
---

```{r load/save image file}
rm(list = ls())
library(ggsignif); library(sjPlot); library(sjmisc); library(sjlabelled); library(modelsummary); source("utils.r"); source("info.r")
encoding <- fread('encoding.csv')
```

Make new variables
```{r}
# re-level
encoding$trialTypeF <- relevel(as.factor(encoding$trialTypeF), ref = "nonliving")
encoding$trialTypeI <- relevel(as.factor(encoding$trialTypeI), ref = "living")
encoding$groupA <- relevel(as.factor(encoding$groupA), ref = 'Adults')
encoding$groupC <- relevel(as.factor(encoding$groupC), ref = 'Children')
```

#################################################
Developmental differences in sustained attention
#############################F####################

Age group difference in lapse rates/time spent out of the zone 
```{r age differences in lapse rates/time out of the zone}
# aggregate data
percents <- encoding[!participant %in% c(outlier$participant), .(percentOut = mean(zone, na.rm=T)), by = .(participant,group, groupEC)]

# model
percents[, summaryh(t.test(percentOut ~ groupEC))]
percents[, percentOutRO := outliersZ(percentOut, zCutOff = 3, replaceOutliersWith = NA, digits= 5), by = group]

# figure 2E
ggplot(percents, aes(group, percentOut *100, fill = as.factor(group))) +
  geom_violin(trim = F, width = 1) +
  geom_quasirandom(width  = .2, size =1 ) +
  ggtheme +
  labs(x = "Mean accuracy", y = "Attentional lapse scores \n (Time out of the zone %)") +
  scale_fill_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) +
  coord_cartesian(ylim = c(-15, 110)) + 
  scale_y_continuous(breaks = c(0, 20, 40, 60, 80, 100)) +
  theme(legend.position = 0) 

```

Relationship between lapse rates and classification accuracy (initial validation reported in methods)
```{r lapse rate and classification accuracy association for children and adults}
mean_accuracy_group <- encoding[, .(acc = mean(acc)), by = .(participant, group, groupEC, groupA, groupC)]
percents <- left_join(percents, mean_accuracy_group)

acc_lapse_rate.m <- lm(acc ~ percentOutRO, data = percents[])
summaryh(acc_lapse_rate.m)
```

Age group difference in lapse frequency
```{r age differences in lapse frequency}
# aggregate data
numOutChunks <- encoding[!participant %in% c(outlier$participant),.(Nolapses = max(chunkNo_out, na.rm=TRUE)), by = .(participant, group, groupEC)]

# model
summaryh(t.test(Nolapses ~ groupEC, data = numOutChunks))

# figure S1
ggplot(numOutChunks[!participant %in% c(outlier$participant)], aes(group, Nolapses, fill = as.factor(group))) +
  geom_violin(trim = F, width = 1) +
  geom_quasirandom(width  = .2, size =1 ) +
  ggtheme +
  labs(x = "Mean accuracy", y = "No. of attentional lapses") +
  scale_fill_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) +
  theme(legend.position = 0)
```

Age group differences in lapse length
```{r age differences in lapse length}
# aggregate data
length_attentional_lapses = encoding[zone == 1 & !participant %in% c(outlier$participant), .N, by = .(participant, group, groupEC, chunkNo)]
length_attentional_lapses2 = encoding[zone == 1 & !participant %in% c(outlier$participant), list(lengthChunk = max(numInChunk)), by = .(participant, group, groupEC, chunkNo)]
#aggregate data
median_length_lapse = length_attentional_lapses[, .(median_lapse_length = median(N)), by = .(participant, group, groupEC)]
#model - pre-registered; note results don't differ depending on model I run
median_length_lapse[!participant %in% c(outlier$participant), summaryh(t.test(median_lapse_length ~ groupEC))]


# figure S2
ggplot(median_length_lapse, aes(group, median_lapse_length, fill = as.factor(group))) +
  geom_violin(trim = F, width = 1) +
  geom_quasirandom(width  = .2, size =1 ) +
  ggtheme +
  labs(x = "", y = "Median lapse length") +
  scale_fill_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) +
  theme(legend.position = 0)
```

Figure 2A,C- Representative child's raw RT and RT deviance
Figure 2A and 2C
```{r  eval = FALSE, echo = FALSE, include =FALSE}
meanrt <- encoding[participant == 224, mean(rt, na.rm=T)]
cutOff <- encoding[, median(smoothed_gaussian, na.rm=TRUE), by = .(participant)][, mean(V1)]

# Figure 2A - raw RT
ggplot(encoding[participant == 224], aes(trialNo, rtResidual + meanrt)) + 
  geom_point(color = 'hotpink3', size = 2) + 
  labs(x = 'Trial number', y = 'Response time (seconds)') + 
  ggtheme +
  coord_cartesian(ylim = c(0, 1.5))

# Figure 2C - RT deviance with zone assignment
toPlot <- encoding[participant == 224 & trialNo %in% c(1:200), .(trialNo, smoothed_gaussian, rtMeanAbsoluteDeviation)]
smoothed_values <- approx(toPlot$trialNo, toPlot$smoothed_gaussian, n = 3500)
non_smoothed_values <- approx(toPlot$trialNo, toPlot$rtMeanAbsoluteDeviation, n = 3500)
toPlot_smoothed <- data.table(trialNo = smoothed_values$x, smoothed_deviance = smoothed_values$y)
toPlot_unsmoothed <- data.table(trialNo = non_smoothed_values$x, unsmoothed_deviance = non_smoothed_values$y)
toPlot2 <- left_join(toPlot_unsmoothed, toPlot_smoothed)
toPlot2[, zone := ifelse(smoothed_deviance > cutOff, 'Out of the zone', 'In the zone')]
encoding[acc == 0, error := 0]
toPlotErrors <- encoding[participant == 224 & trialNo %in% c(1:200), .(trialNo, error)]

ggplot(toPlot2, aes(trialNo, smoothed_deviance, col = zone)) + 
  geom_line(data = toPlot2, aes(trialNo, y = unsmoothed_deviance), inherit.aes = F, color = 'lightgrey') +
  geom_path(group = 1, size = 1.5) +
  labs(x  = "Trial Number", y = "Smoothed RT deviance") +
  ggtheme + 
  scale_color_manual(breaks = c("In the zone", "Out of the zone"), values = c("maroon4", "hotpink1")) + 
  geom_point(data = toPlotErrors, aes(trialNo, y = error), inherit.aes = F, color = 'red', shape = 15) + 
  scale_y_continuous(breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6))+
  coord_cartesian(ylim = c(0, 0.6)) + 
  theme(legend.position = 0)
```

Figure 2B, D - Representative adult's raw RT and RT deviance
```{r Figure 2B and 2D eval = FALSE, echo = FALSE, include =FALSE}
# Figue 2B
meanrt <- encoding[participant == 157, mean(rt, na.rm=T)]
ggplot(encoding[participant == 157], aes(trialNo, rtResidual + meanrt)) + 
  geom_point(color = 'royalblue3', size = 2) + 
  labs(x = 'Trial Number', y = 'Response time (seconds)') + 
  ggtheme +
  coord_cartesian(ylim = c(0, 1.5))

# Figure 2D
toPlot <- encoding[participant == 157 & trialNo %in% c(1:200), .(trialNo, smoothed_gaussian, rtMeanAbsoluteDeviation)]
smoothed_values <- approx(toPlot$trialNo, toPlot$smoothed_gaussian, n = 3500)
non_smoothed_values <- approx(toPlot$trialNo, toPlot$rtMeanAbsoluteDeviation, n = 3500)
toPlot_smoothed <- data.table(trialNo = smoothed_values$x, smoothed_deviance = smoothed_values$y)
toPlot_unsmoothed <- data.table(trialNo = non_smoothed_values$x, unsmoothed_deviance = non_smoothed_values$y)
toPlot2 <- left_join(toPlot_unsmoothed, toPlot_smoothed)
toPlot2[, zone := ifelse(smoothed_deviance > cutOff, 'Out of the zone', 'In the zone')]
encoding[acc == 0, error := 0]
toPlotErrors <- encoding[participant == 157 & trialNo %in% c(1:200), .(trialNo, error)]

# Figure 2D
ggplot(toPlot2, aes(trialNo, smoothed_deviance, col = zone)) + 
  geom_line(data = toPlot2, aes(trialNo, y = unsmoothed_deviance), inherit.aes = F, color = 'lightgrey') +
  geom_path(group = 1, size = 1.5) +
  labs(x  = "Trial Number", y = "Smoothed RT deviance") +
  ggtheme + 
  scale_color_manual(breaks = c("In the zone", "Out of the zone"), values = c("royalblue4", "steelblue1")) + 
  geom_point(data = toPlotErrors, aes(trialNo, y = error), inherit.aes = F, color = 'red', shape = 15) + 
  scale_y_continuous(breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6))+
  coord_cartesian(ylim = c(0, 0.6)) + 
  theme(legend.position = 0)
```

#########################################################
Developmental differences in classification performance
#########################################################

```{r Age group differences in classification performance}
# aggregate data
aggregate_accuracy_group <- encoding[, .(acc = mean(acc)), by = .(participant, group, groupEC, groupA, groupC, trialTypeEC, trialType, trialTypeI, trialTypeF)]
encoding$groupEC
accuracy_group.m_agg <- lmer(acc~groupEC * trialTypeEC + (1|participant), data = aggregate_accuracy_group[!participant %in% c(outlier$participant)]);summaryh(accuracy_group.m_agg)

# separately for children and adults:
aggregate_accuracy_group[!participant %in% c(outlier$participant), summaryh(lm(acc ~ groupEC)), by = trialTypeEC][term != "(Intercept)"]
encoding[, summaryh(lmer(acc ~ groupC * trialTypeEC + (1|participant)))]
encoding[, summaryh(glmer(acc ~ groupC * trialTypeEC + (1|participant), family = "binomial"))]


ggplot(aggregate_accuracy_group, aes(trialType, acc, col = group)) + 
         stat_summary()

```

Between subjects: Accuracy and Attention relationships 
```{r relationship between attentional lapse rates and classification accuracy}
# model: accuracy and attention correlation
percents_trialtype <- left_join(aggregate_accuracy_group[!participant %in% c(outlier$participant)], percents[, .(participant, percentOut)])
percents_trialtype[, percentOutC := scale(percentOut, scale = F)]

percentOut_accuracy.m <- lmer(acc ~ percentOutC * trialTypeEC * groupEC + (1|participant), data = percents_trialtype[!participant %in% c(outlier$participant)]); summaryh(percentOut_accuracy.m)

percents_trialtype[!participant %in% c(outlier$participant), summaryh(lm(acc ~ percentOutC)), by = .(group, trialTypeEC)][term != "(Intercept)"]
percents_trialtype[!participant %in% c(outlier$participant), summaryh(lm(acc ~ percentOutC)), by = .(group, trialTypeEC)][term != "(Intercept)"]


# Visualize - accuracy and attention correlation
ggplot(percents[!participant %in% c(outlier$participant)], aes(percentOut, acc, col = group))+
  geom_point() +
  ggtheme+
  stat_smooth(formula = y ~ x, method = "lm") +
  labs(x = "Time out of the zone (%)", y = "Mean accuracy (%)") +
  scale_x_continuous(breaks = c(0, .25, .50, .75), labels = c(0, 25, 50, 75)) +
  scale_y_continuous(breaks = c(.9, .95, 1), labels = c(90, 95, 100)) +
scale_color_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) +
    coord_cartesian(ylim = c(.9, 1.01))+
  theme(legend.position = 0)

# Figure S4A
ggplot(percents_trialtype[!participant %in% c(outlier$participant) & trialType == 'nonliving'], aes(percentOutC, acc, col = group))+
  geom_point() +
  ggtheme +
  stat_smooth(formula = y ~ x, method = "lm") +
  labs(x = "Time out of the zone (%)", y = "Mean accuracy (%)") +
  scale_x_continuous(breaks = c(0, .25, .50, .75), labels = c(0, 25, 50, 75)) +
  scale_y_continuous(breaks = c(.94, .96, .98, 1), labels = c(94, 96, 98, 100)) +
scale_color_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) + 
  theme(legend.position = 0)

# Figure S4B
ggplot(percents_trialtype[!participant %in% c(outlier$participant) & trialType == 'living'], aes(percentOut, acc, col = group))+
  geom_point() +
  ggtheme +
  stat_smooth(formula = y ~ x, method = "lm") +
  labs(x = "Time out of the zone (%)", y = "Mean accuracy (%)") +
  scale_x_continuous(breaks = c(0, .25, .50, .75), labels = c(0, 25, 50, 75)) +
  scale_y_continuous(breaks = c(.4, .6, .8, 1), labels = c(40, 60, 80, 100)) +
  coord_cartesian(ylim = c(.42, 1.018), xlim = c(0.05, .85))+
scale_color_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) + 
  theme(legend.position = 0)
```

Re-do above analyses comparing children to matched adults
Check whether attentional lapse rates shape classication in high vs low lapsing adults
```{r}
### two groups of adults ###
library(stats)
dataToCluster <- percents[group == "Adults" & !participant %in% c(outlier$participant), .(participant, percentOut)]
dataToCluster[, percentOutZ := scale(percentOut)]
dataToCluster$kmeansGroup <- kmeans(dataToCluster$percentOutZ, centers = 2, iter.max = 1000, algorithm = "MacQueen")[1]
dataToCluster[, .N, by = kmeansGroup]
dataToCluster[, mean(percentOutZ), by = kmeansGroup]
colors <- c(colors, "mediumorchid4", 'yellow4')
dataToCluster <- rbind(dataToCluster, percents[group == "Children" & !participant %in% c(outlier$participant), .(participant, percentOut)], fill = T)

# join data
percents <- left_join(percents, dataToCluster)
percents[, .N, by = .(kmeansGroup)]
percents[, percentOutC := scale(percentOut, scale = F)]
# data
percents_trialtype <- left_join(aggregate_accuracy_group[!participant %in% c(outlier$participant)], percents[, .(participant, percentOut, percentOutC, kmeansGroup)])
```

```{r}
# do lapses influence accuracy more in children than high lapsing adults (separately by trial type) 
percents_trialtype[!kmeansGroup %in% c(1), summaryh(lm(acc ~ percentOutC * groupEC)), by = .( trialTypeEC)][!term %in% "(Intercept)"]

# do lapses influence accuracy more in children than low lapsing adults (separately by trial type) 
percents_trialtype[!kmeansGroup %in% c(2), summaryh(lm(acc ~ percentOutC * groupEC)), by = .( trialTypeEC)][!term %in% "(Intercept)"]

percents_trialtype[kmeansGroup == 1, kmeansGroupEC := -0.5]
percents_trialtype[kmeansGroup == 2, kmeansGroupEC := 0.5]
percents_trialtype[!kmeansGroup %in% c(3), summaryh(lm(acc ~ percentOutC * kmeansGroupEC)), by = .( trialTypeEC)][!term %in% "(Intercept)"]

# low vs children
percents_trialtype[!participant %in% c(outlier$participant) & !kmeansGroupEC %in% c(1), 
                   summaryh(lm(acc ~ percentOutC * groupEC)), by = .(trialType)]
percents[!participant %in% c(outlier$participant) & !kmeansGroup %in% c(1), 
                   summaryh(lm(acc ~ percentOutC * groupEC))]


# high vs children
percents_trialtype[!participant %in% c(outlier$participant) & !kmeansGroupEC %in% c(-1), 
                   summaryh(lm(acc ~ percentOutC * groupEC)), by = .(trialType)]
percents[!participant %in% c(outlier$participant) & !kmeansGroup %in% c(2), 
                   summaryh(lm(acc ~ percentOutC * groupEC))]


# within high vs. low lapsing adults
percents[!participant %in% c(outlier$participant), 
                   summaryh(lm(acc ~ percentOutC)), by = .(kmeansGroup)][term !="(Intercept)"]
percents_trialtype[!participant %in% c(outlier$participant), 
                   summaryh(lm(acc ~ percentOutC)), by = .(trialType, kmeansGroup)][term !="(Intercept)"]
percents[!participant %in% c(outlier$participant), mean(percentOut), by = kmeansGroup]

percents_trialtype[, kmeansGroupEC := ifelse(kmeansGroup == 1, -0.5, 0.5)]
```

```{r figure relating percent out to classification separately for high vs. low lapsing adults}
ggplot(percents[!participant %in% c(outlier$participant) & group %in% ("Adults")], aes(percentOut*100, acc*100, col = as.factor(kmeansGroup)))+
  geom_point() +
  ggtheme+
  stat_smooth(data = percents[kmeansGroup == 1], formula = y ~ x, method = "lm") +
  stat_smooth(data = percents[kmeansGroup == 2], formula = y ~ x, method = "lm") +
  labs(x = "Lapse rates (%)", y = "Classification accuracy (%)") +
  scale_color_manual(values = c("blue4", "royalblue1")) +
  theme(legend.position = 0) +
  coord_cartesian(ylim = c(95, 100)) + 
  scale_y_continuous(breaks = c(94, 96, 98, 100), labels = c("94", "96", "98", "100"))
```

#### Accuracy and time out mediation ####
```{r eval = FALSE, include = FALSE, echo = FALSE}
#acme: mediated (ab) effect
#ade: direct effect (c')
#total effect: c
percents[!participant %in% c(outlier$participant), percentOutZ := scale(percentOut)]
percents[!participant %in% c(outlier$participant), mean_acc.allZ := scale(acc)]

model.y_all = lm(mean_acc.allZ ~ groupEC + percentOutZ , data = percents[!participant %in% c(outlier$participant)])
summary(model.y_all)

model.m_all = lm(percentOutZ ~ groupEC, data = percents[!participant %in% c(outlier$participant)])
summary(model.m_all)

model.mediate_attention_accuracy_all = mediate(model.m_all, model.y_all, sims = 5000, treat = 'groupEC', mediator = 'percentOutZ', covariates = 'mean_rt', boot = TRUE)
summary(model.mediate_attention_accuracy_all)
```


#################################################################
Within subject analysis predicting accuracy from RT deviance
#################################################################

model for accuracy ~ trial type * group * age

```{r eval = FALSE, include = FALSE, echo = FALSE}
# model: Age differences
continuous_attention_accuracy_age_interaction.m = glmer(acc ~ pre_smoothed_deviance_scaled * trialTypeEC * groupEC + (pre_smoothed_deviance_scaled * trialTypeEC||participant), control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)), family  = "binomial", data = encoding); summaryh(continuous_attention_accuracy_age_interaction.m)

# Visualize overall relationship between accuracy and RT deviance
toPlot <- encoding[!is.na(acc) & !is.na(pre_smoothed_deviance), .(rt_deviance = mean(pre_smoothed_deviance_scaled, na.rm=T)), by = .(participant, group, acc)]
toPlot.w <- seWithin(data = toPlot, measurevar = "rt_deviance", betweenvars = "group", withinvars = "acc", idvar = "participant")

ggplot(toPlot.w, aes(as.factor(group), rt_deviance, col = as.factor(acc))) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_errorbar(aes(ymin = rt_deviance - se, ymax = rt_deviance + se), position = position_dodge(width = 0.5), width = 0, size = 1) + 
  ggtheme + 
#scale_color_viridis(discrete = TRUE, option = "G", labels = c("Error", "Correct"))  +
  scale_color_manual(breaks = c(0, 1), labels = c("Error", "Correct"), values = c("orchid4", "seagreen4")) +
  labs(y = "Preceding RT deviance (s)", x = " ") 
```



```{r }
# re-do graph so it's easier to interpret (with memory hits on the y axis, age group on the x axis, and preceding deviance on the color)
#toPlot <- encoding[!is.na(acc) & !is.na(pre_smoothed_deviance), .(rt_deviance = mean(pre_smoothed_deviance_scaled, na.rm=T)), by = .(participant, group, acc)]
encoding[, median_pre_smoothed_deviance := median(pre_smoothed_deviance, na.rm=T), by = participant]
encoding[, pre_smoothed_deviance_cat := ifelse(pre_smoothed_deviance > median_pre_smoothed_deviance, "high", "low")]
encoding[pre_smoothed_deviance == median_pre_smoothed_deviance, pre_smoothed_deviance_cat := "low"]
encoding[, mean(pre_smoothed_deviance), by = .(pre_smoothed_deviance_cat)]
encoding[is.na(pre_smoothed_deviance_cat) & !is.na(pre_smoothed_deviance)]

toPlot <- encoding[!is.na(pre_smoothed_deviance_cat), .(mean_acc = mean(acc, na.rm=T)), by = .(participant, group, pre_smoothed_deviance_cat)]
toPlot.w <- seWithin(data = toPlot, measurevar = "mean_acc", betweenvars = "group", withinvars = "pre_smoothed_deviance_cat", idvar = "participant")

ggplot(toPlot.w, aes(group, mean_acc*100, col = pre_smoothed_deviance_cat)) +
  #stat_summary()+
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_errorbar(aes(ymin = mean_acc*100 - se*100, ymax = mean_acc*100 + se*100), position = position_dodge(width = 0.5), width = 0, size = 1) + 
  ggtheme  + 
#scale_color_viridis(discrete = TRUE, option = "G", labels = c("Error", "Correct"))  +
  scale_color_manual(breaks = c("low", "high"), labels = c("low deviance", "high deviance"), values = c("orchid4", "seagreen4")) +
  scale_y_continuous(breaks = c(94, 96, 98, 100), labels = c("94", "96", "98", "100"))+
  coord_cartesian(ylim = c(94, 100)) +
  labs(y = "Accuracy (%)", x = " ", col = " ") 
```


######################
Memory analysis #####
######################

Group differences in d prime high confidence 
```{r calc dprime}
recognition <- fread(file = 'recognition.csv')
recognition[, groupEC := ifelse(group == "Children", -0.5, 0.5)]
recognition[, trialTypeEC := ifelse(classifyEnc == "living", -0.5, 0.5)]

# age group variables
recognition[, groupC := group]
recognition[, groupA := group]
recognition$groupA <- relevel(as.factor(recognition$groupA), ref = "Adults")
recognition$groupC <- relevel(as.factor(recognition$groupC), ref = "Children")

# trial type variable
#recognition[, trialTypeEC := classifyEncEC]
recognition[, trialTypeI := classifyEnc]
recognition[, trialTypeF := classifyEnc]
recognition$trialTypeI <- relevel(as.factor(recognition$trialTypeI), ref = "living")
recognition$trialTypeF <- relevel(as.factor(recognition$trialTypeF), ref = "nonliving")

#overall d prime
recognition[, memory_hits := memoryHits_dprime]
recognition[classifyMemory == "old" & choiceMemory == "new", memory_hits := 0]
recognition[ratingConfidence %in% c(1, 2), .(choiceMemory, classifyMemory,ratingConfidence, acc, memoryHits_dprime, memory_hits, false_alarm_dprime, false_alarm_dprime_all_conf)]

# changed from hits = memoryHits_dprime to memory_hits; changed false_alarms from false_alarm_dprime to false_alarm_dprime_all_conf
Dprime.hc <- recognition[, dprime(hits = memoryHits_dprime, false_alarms = false_alarm_dprime, na.rm = T), by = .(participant, group, groupEC, groupA, groupC)]

#by trial type
dprime.hc_trialtype <- recognition[, dprime(hits = memoryHits_dprime, false_alarms = false_alarm_dprime, na.rm = T), by = .(participant, group, groupEC, groupA, groupC, trialTypeI, trialTypeF, trialTypeEC)]
```

```{r dprime by trial type}
# frequent vs. infrequent trials
dprime.hc_trialtype[, dprimeC := scale(dprime, scale = F)]
dprime.hc_trialtype[!participant %in% c(outlier$participant), summaryh(lmer(dprime ~ groupEC * trialTypeEC + (1|participant)))][term != '(Intercept)']
dprime.hc_trialtype[!participant %in% c(outlier$participant), summaryh(lm(dprime ~ trialTypeEC + (1|participant))), by = group][term != '(Intercept)']
```

```{r dprime figure}
# summary
# overall
ggplot(Dprime.hc[!participant %in% c(outlier$participant)], aes(group, dprime, fill = as.factor(group))) +
  geom_violin(trim = F) +
      geom_quasirandom(width  = .2, size = 1) +
  ggtheme +
  labs(x = "", y = "D prime") +
scale_fill_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) +
  theme(legend.position = 0) + 
  ylim(0.5,4)

# nonliving trials
ggplot(dprime.hc_trialtype[!participant %in% c(outlier$participant) & trialTypeF == 'nonliving'], aes(group, dprime, fill = as.factor(group))) +
  geom_violin(trim = F) +
      geom_quasirandom(width  = .2, size = 1) +
  ggtheme +
  labs(x = "", y = "D prime") +
scale_fill_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) +
  theme(legend.position = 0) + 
  ylim(0.5,4)

# living trials
ggplot(dprime.hc_trialtype[!participant %in% c(outlier$participant) & trialTypeI == 'living'], aes(group, dprime, fill = as.factor(group))) +
  geom_violin(trim = F) +
      geom_quasirandom(width  = .2, size = 1) +
  ggtheme +
  labs(x = "", y = "D prime") +
scale_fill_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) +
  theme(legend.position = 0) +
    ylim(0.15,4.5)
```

```{r dprime figure by age and trial type}
ggplot(dprime.hc_trialtype, aes(group, dprime, col = group)) + 
  #geom_violin(trim = F) +
      #geom_quasirandom(width  = .2, size = 1) +
  ggtheme +
  labs(x = "", y = "d prime") +
scale_color_manual(values = c("darkblue", "mediumvioletred")) +
  stat_summary()  + 
  facet_wrap(~trialTypeF) 
```

#########################################################
Correlations between d prime and time out of the zone 
########################################################

Join data
```{r join dprime to time out data}
# join variables to overall dprime dataframe
percents[, percentOutC := scale(percentOut, scale=F)]

percents_dprime.hc = left_join(Dprime.hc, percents[, .N, by = .(participant, percentOut, percentOutC, groupEC)][, .(participant, percentOut, percentOutC)])
percents_dprime.hc = left_join(percents_dprime.hc, numOutChunks)
percents_dprime.hc = left_join(percents_dprime.hc, median_length_lapse)

# join variables to dataframe by trial types
percents_dprime.hc_trialtype = left_join(dprime.hc_trialtype, percents[, .N, by = .(participant, percentOut, percentOutC, groupEC)][, .(participant, percentOut, percentOutC)])
percents_dprime.hc_trialtype = left_join(percents_dprime.hc_trialtype, numOutChunks)
percents_dprime.hc_trialtype = left_join(percents_dprime.hc_trialtype, median_length_lapse)
```

Run correlations
```{r overall - collapsed across trial types}
# dprime correlations with percent out

percents_dprime.hc[!participant %in% c(outlier$participant), summaryh(lm(dprime ~ percentOutC*groupEC))][!term == "(Intercept)"]
percents_dprime.hc[!participant %in% c(outlier$participant), summaryh(lm(dprime ~ percentOutC)), by = group][!term == "(Intercept)"]
```

```{r figure attentional lapse rate and memory relationships by group}
ggplot(percents_dprime.hc[!participant %in% c(outlier$participant)], aes(percentOut *100, dprime, col = group))+
  geom_point() +
  ggtheme+
  stat_smooth(formula = y ~ x, method = "lm") +
  labs(x = "Time out of the zone (%)", y = "D prime") +
  #scale_x_continuous(breaks = c(0, .2, .4, .6, 0.8), labels = c(0, 20, 40, 60, 80)) +
  scale_y_continuous(breaks = c(1, 2, 3), labels = c(1, 2, 3)) +
  #coord_cartesian(ylim = c(1, 3), xlim = c(0, .8)) +
  scale_color_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) +
  theme(legend.position = 0)


ggplot(percents_dprime.hc_trialtype[!participant %in% c(outlier$participant) & trialTypeEC == 0.5], aes(percentOut *100, dprime, col = group))+
  geom_point() +
  ggtheme+
  stat_smooth(formula = y ~ x, method = "lm") +
  labs(x = "Lapse rate (%)", y = "D prime") +
  #scale_x_continuous(breaks = c(0, .2, .4, .6, 0.8), labels = c(0, 20, 40, 60, 80)) +
  scale_y_continuous(breaks = c(1, 2, 3), labels = c(1, 2, 3)) +
  #coord_cartesian(ylim = c(1, 3), xlim = c(0, .8)) +
  scale_color_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) +
  theme(legend.position = 0) #+ 
  #coord_cartesian(ylim = c(0.5,3.5))


# infrequent
ggplot(percents_dprime.hc_trialtype[!participant %in% c(outlier$participant) & trialTypeEC == -0.5], aes(percentOut *100, dprime, col = group))+
  geom_point() +
  ggtheme+
  stat_smooth(formula = y ~ x, method = "lm") +
  labs(x = "Lapse rate (%)", y = "D prime") +
  #scale_x_continuous(breaks = c(0, .2, .4, .6, 0.8), labels = c(0, 20, 40, 60, 80)) +
  scale_y_continuous(breaks = c(1, 2, 3), labels = c(1, 2, 3)) +
  #coord_cartesian(ylim = c(1, 3), xlim = c(0, .8)) +
  scale_color_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) +
  theme(legend.position = 0)
```


###############################
Within subject memory analyses 
##############################

Relating continuous measures of attention to memory hits
```{r wrangle data}
recognition[classifyMemory == 'old' & choiceMemory == 'old'  & ratingConfidence %in% c(3, 4), memory_hits := 1]
recognition[classifyMemory == 'old' & choiceMemory == 'new', memory_hits := 0]
encoding$memory_hits <- NULL
encoding = left_join(encoding, recognition[classifyMemory == 'old', .(participant, stimulus, memory_hits)])
encoding <- left_join(encoding, median_length_lapse)
encoding <- left_join(encoding, numOutChunks)
```

```{r models}
# rt preceding deviance
fluctuations_hits.m <- encoding[, summaryh(glmer(memory_hits ~ pre_smoothed_deviance_scaled  * classifyEncEC * groupEC +  (pre_smoothed_deviance_scaled * classifyEncEC ||participant), control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)), family = 'binomial'))][term != '(Intercept)']
```

```{r figure - overall memory}
smoothed_deviance_memory_hits <- encoding[!is.na(memory_hits), .(prec_rt_dev = mean(pre_smoothed_deviance_scaled, na.rm=T)), by = .(participant, group, memory_hits)]

deviance_memory_prep.p = seWithin(smoothed_deviance_memory_hits, measurevar = "prec_rt_dev", betweenvars = c("group"), withinvars = c("memory_hits"), idvar = "participant")

ggplot(deviance_memory_prep.p[!is.na(memory_hits)], aes(as.factor(memory_hits), prec_rt_dev, col = group)) +
   stat_summary(fun.y = "mean", geom = "point", position = position_dodge(width = 0.5), size  = 3) +
  geom_errorbar(aes(ymin = prec_rt_dev - se, ymax = prec_rt_dev + se), position = position_dodge(width = 0.5), width = 0, size = 1)+
    labs(y = "Preceding RT deviance", x = "Memory hits", col = " ")+
  ggtheme +
scale_color_manual(values = c("royalblue3", "hotpink3")) +
  scale_x_discrete(labels = c("Forgotten", "Remembered"))
```
