---
output: html_document
editor_options: 
  chunk_output_type: console
---
--
title: "R Notebook"
output: html_notebook
chunk_output_type: console
---

```{r load data and libraries}
rm(list = ls())    
library(ggsignif); library(sjPlot); library(sjmisc); library(sjlabelled); library(modelsummary); source("utils.r"); source("info.r"); library(stats)
encoding <- fread('../data/encoding.csv')
```

The code below is reported under the heading "Children's attention lapsed twice as frequently as adults'"

The chunk below computes age group difference in time spent out of the zone (lapse rates).
```{r age differences in lapse rates/time out of the zone}
# calculate the percent of time each participant spends out of the zone
percents <- encoding[!participant %in% c(outlier$participant), .(percentOut = mean(zone, na.rm=T)), by = .(participant,group, groupEC)]

# fit model to test for age group differences in time spent out of the zone/lapse rates
percents[, summaryh(t.test(percentOut ~ groupEC))]

# plot figure 2E
ggplot(percents, aes(group, percentOut *100, fill = as.factor(group))) +
  geom_violin(trim = F, width = 1) +
  geom_quasirandom(width  = .2, size =1 ) +
  ggtheme +
  labs(x = "Mean accuracy", y = "Attentional lapse scores \n (Time out of the zone %)") +
  scale_fill_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) +
  coord_cartesian(ylim = c(-15, 110)) + 
  scale_y_continuous(breaks = c(0, 20, 40, 60, 80, 100)) +
  theme(legend.position = "None")
```

The chunk belwo computes age group differences in lapse length
```{r age differences in lapse length}
# compute lapse length per participant
# it selects zone ==1 meaning time when people are "out" of the zone
length_attentional_lapses = encoding[zone == 1 & !participant %in% c(outlier$participant), .N, by = .(participant, group, groupEC, chunkNo)]

# compute a datatable that contains the median lapse length per participant
median_length_lapse = length_attentional_lapses[, .(median_lapse_length = median(N)), by = .(participant, group, groupEC)]

#model - pre-registered
median_length_lapse[!participant %in% c(outlier$participant), summaryh(t.test(median_lapse_length ~ groupEC))]

# figure S2
ggplot(median_length_lapse, aes(group, median_lapse_length, fill = as.factor(group))) +
  geom_violin(trim = F, width = 1) +
  geom_quasirandom(width  = .2, size =1 ) +
  ggtheme +
  labs(x = "", y = "Median lapse length") +
  scale_fill_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) +
  theme(legend.position = "None")
```

Age group difference in lapse frequency
```{r age differences in lapse frequency}
# compute the lapse frequency
numOutChunks <- encoding[!participant %in% c(outlier$participant),.(Nolapses = max(chunkNo_out, na.rm=TRUE)), by = .(participant, group, groupEC)]

# model
summaryh(t.test(Nolapses ~ groupEC, data = numOutChunks))

# figure S1
ggplot(numOutChunks[!participant %in% c(outlier$participant)], aes(group, Nolapses, fill = as.factor(group))) +
  geom_violin(trim = F, width = 1) +
  geom_quasirandom(width  = .2, size =1 ) +
  ggtheme +
  labs(x = "Mean accuracy", y = "No. of attentional lapses") +
  scale_fill_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) +
  theme(legend.position = "None")
```

The code below plots Figure 2A,C- Representative child's raw RT and RT deviance
Figure 2A and 2C
```{r  eval = FALSE, echo = FALSE, include =FALSE}
meanrt <- encoding[participant == 224, mean(rt, na.rm=T)]
cutOff <- encoding[, median(smoothed_gaussian, na.rm=TRUE), by = .(participant)][, mean(V1)]

# Figure 2A - raw RT
ggplot(encoding[participant == 224], aes(trialNo, rtResidual + meanrt)) + 
  geom_point(color = 'hotpink3', size = 2) + 
  labs(x = 'Trial number', y = 'Response time (seconds)') + 
  ggtheme +
  coord_cartesian(ylim = c(0, 1.5))

# Figure 2C - RT deviance with zone assignment
toPlot <- encoding[participant == 224 & trialNo %in% c(1:200), .(trialNo, smoothed_gaussian, rtMeanAbsoluteDeviation)]
smoothed_values <- approx(toPlot$trialNo, toPlot$smoothed_gaussian, n = 3500)
non_smoothed_values <- approx(toPlot$trialNo, toPlot$rtMeanAbsoluteDeviation, n = 3500)
toPlot_smoothed <- data.table(trialNo = smoothed_values$x, smoothed_deviance = smoothed_values$y)
toPlot_unsmoothed <- data.table(trialNo = non_smoothed_values$x, unsmoothed_deviance = non_smoothed_values$y)
toPlot2 <- left_join(toPlot_unsmoothed, toPlot_smoothed)
toPlot2[, zone := ifelse(smoothed_deviance > cutOff, 'Out of the zone', 'In the zone')]
encoding[acc == 0, error := 0]
toPlotErrors <- encoding[participant == 224 & trialNo %in% c(1:200), .(trialNo, error)]

ggplot(toPlot2, aes(trialNo, smoothed_deviance, col = zone)) + 
  geom_line(data = toPlot2, aes(trialNo, y = unsmoothed_deviance), inherit.aes = F, color = 'lightgrey') +
  geom_path(group = 1, size = 1.5) +
  labs(x  = "Trial Number", y = "Smoothed RT deviance") +
  ggtheme + 
  scale_color_manual(breaks = c("In the zone", "Out of the zone"), values = c("maroon4", "hotpink1")) + 
  geom_point(data = toPlotErrors, aes(trialNo, y = error), inherit.aes = F, color = 'red', shape = 15) + 
  scale_y_continuous(breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6))+
  coord_cartesian(ylim = c(0, 0.6)) + 
  theme(legend.position = "None")
```

The code below plots Figure 2B, D - Representative adult's raw RT and RT deviance
```{r Figure 2B and 2D eval = FALSE, echo = FALSE, include =FALSE}
# Figue 2B
meanrt <- encoding[participant == 157, mean(rt, na.rm=T)]
ggplot(encoding[participant == 157], aes(trialNo, rtResidual + meanrt)) + 
  geom_point(color = 'royalblue3', size = 2) + 
  labs(x = 'Trial Number', y = 'Response time (seconds)') + 
  ggtheme +
  coord_cartesian(ylim = c(0, 1.5))

# Figure 2D
toPlot <- encoding[participant == 157 & trialNo %in% c(1:200), .(trialNo, smoothed_gaussian, rtMeanAbsoluteDeviation)]
smoothed_values <- approx(toPlot$trialNo, toPlot$smoothed_gaussian, n = 3500)
non_smoothed_values <- approx(toPlot$trialNo, toPlot$rtMeanAbsoluteDeviation, n = 3500)
toPlot_smoothed <- data.table(trialNo = smoothed_values$x, smoothed_deviance = smoothed_values$y)
toPlot_unsmoothed <- data.table(trialNo = non_smoothed_values$x, unsmoothed_deviance = non_smoothed_values$y)
toPlot2 <- left_join(toPlot_unsmoothed, toPlot_smoothed)
toPlot2[, zone := ifelse(smoothed_deviance > cutOff, 'Out of the zone', 'In the zone')]
encoding[acc == 0, error := 0]
toPlotErrors <- encoding[participant == 157 & trialNo %in% c(1:200), .(trialNo, error)]

# Figure 2D
ggplot(toPlot2, aes(trialNo, smoothed_deviance, col = zone)) + 
  geom_line(data = toPlot2, aes(trialNo, y = unsmoothed_deviance), inherit.aes = F, color = 'lightgrey') +
  geom_path(group = 1, size = 1.5) +
  labs(x  = "Trial Number", y = "Smoothed RT deviance") +
  ggtheme + 
  scale_color_manual(breaks = c("In the zone", "Out of the zone"), values = c("royalblue4", "steelblue1")) + 
  geom_point(data = toPlotErrors, aes(trialNo, y = error), inherit.aes = F, color = 'red', shape = 15) + 
  scale_y_continuous(breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6))+
  coord_cartesian(ylim = c(0, 0.6)) + 
  theme(legend.position = "None")
```

The following section of code is reported under the heading "Children's higher lapse rates underlie their lower categorization performance"

First, compute age group differences in classification performance on the sustained attention task:
```{r Age group differences in classification performance}
# compute classification accuracy for each participant 
aggregate_accuracy_group <- encoding[, .(acc = mean(acc)), by = .(participant, group, groupEC, groupA, groupC, trialTypeEC, trialType, trialTypeI, trialTypeF)]

# fit model examining age group differences in classification accuracy
accuracy_group.m_agg <- lmer(acc ~ groupEC * trialTypeEC + (1|participant), data = aggregate_accuracy_group[!participant %in% c(outlier$participant)]);summaryh(accuracy_group.m_agg)

# fit model separately for children and adults:
aggregate_accuracy_group[!participant %in% c(outlier$participant), summaryh(lm(acc ~ groupEC)), by = trialTypeEC][term != "(Intercept)"]
encoding[, summaryh(lmer(acc ~ groupC * trialTypeEC + (1|participant)))]
encoding[, summaryh(glmer(acc ~ groupC * trialTypeEC + (1|participant), family = "binomial"))]
```

Then, test how individual differences in lapse rates relate to classification performance:
```{r relationship between attentional lapse rates and classification accuracy}
# join data on lapse rates and accuracy:
percents_trialtype <- left_join(aggregate_accuracy_group[!participant %in% c(outlier$participant)], percents[, .(participant, percentOut)])

# center the lapse rate variable:
percents_trialtype[, percentOutC := scale(percentOut, scale = F)]

# fit model relating time out of the zone to classification accuracy
percentOut_accuracy.m <- lmer(acc ~ percentOutC * trialTypeEC * groupEC + (1|participant), data = percents_trialtype[!participant %in% c(outlier$participant)]); summaryh(percentOut_accuracy.m)

# fit model separately for children and adults:
percents_trialtype[!participant %in% c(outlier$participant), summaryh(lm(acc ~ percentOutC)), by = .(group, trialTypeEC)][term != "(Intercept)"]

# join average accuracy and attention data:
mean_accuracy_group <- encoding[, .(acc = mean(acc)), by = .(participant, groupA, groupC)]
percents <- left_join(percents, mean_accuracy_group)

# Visualize - accuracy and attention correlation:
ggplot(percents[!participant %in% c(outlier$participant)], aes(percentOut, acc, col = group))+
  geom_point() +
  ggtheme+
  stat_smooth(formula = y ~ x, method = "lm") +
  labs(x = "Time out of the zone (%)", y = "Mean accuracy (%)") +
  scale_x_continuous(breaks = c(0, .25, .50, .75), labels = c(0, 25, 50, 75)) +
  scale_y_continuous(breaks = c(.9, .95, 1), labels = c(90, 95, 100)) +
scale_color_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) +
    coord_cartesian(ylim = c(.9, 1.01))+
  theme(legend.position = "None")

# Figure S4A
ggplot(percents_trialtype[!participant %in% c(outlier$participant) & trialType == 'nonliving'], aes(percentOutC, acc, col = group))+
  geom_point() +
  ggtheme +
  stat_smooth(formula = y ~ x, method = "lm") +
  labs(x = "Time out of the zone (%)", y = "Mean accuracy (%)") +
  scale_x_continuous(breaks = c(0, .25, .50, .75), labels = c(0, 25, 50, 75)) +
  scale_y_continuous(breaks = c(.94, .96, .98, 1), labels = c(94, 96, 98, 100)) +
scale_color_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) + 
  theme(legend.position = "None")

# Figure S4B
ggplot(percents_trialtype[!participant %in% c(outlier$participant) & trialType == 'living'], aes(percentOut, acc, col = group))+
  geom_point() +
  ggtheme +
  stat_smooth(formula = y ~ x, method = "lm") +
  labs(x = "Time out of the zone (%)", y = "Mean accuracy (%)") +
  scale_x_continuous(breaks = c(0, .25, .50, .75), labels = c(0, 25, 50, 75)) +
  scale_y_continuous(breaks = c(.4, .6, .8, 1), labels = c(40, 60, 80, 100)) +
  coord_cartesian(ylim = c(.42, 1.018), xlim = c(0.05, .85))+
scale_color_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) + 
  theme(legend.position = "None")
```

Exploratory analysis to test whether lapses relate to accuracy in "high lapsing" adults.
```{r}
### two groups of adults ###
# select a dataset of adults only for classifying adults into those who lapse more vs. less
dataToCluster <- percents[group == "Adults" & !participant %in% c(outlier$participant), .(participant, percentOut)]

# scale the percent of time individuals spent off task:
dataToCluster[, percentOutZ := scale(percentOut)]

# fit k means cluster model:
dataToCluster$kmeansGroup <- kmeans(dataToCluster$percentOutZ, centers = 2, iter.max = 1000, algorithm = "MacQueen")[1]

# bind the group labels to the percents data table:
dataToCluster <- rbind(dataToCluster, percents[group == "Children" & !participant %in% c(outlier$participant), .(participant, percentOut)], fill = T)

# join data
percents <- left_join(percents, dataToCluster)
# scale the percent out data
percents[, percentOutC := scale(percentOut, scale = F)]

# data
percents_trialtype <- left_join(aggregate_accuracy_group[!participant %in% c(outlier$participant)], percents[, .(participant, percentOut, percentOutC, kmeansGroup)])

# fit model relating lapse rates to accuracy in high and low lapsing adults:
percents[!participant %in% c(outlier$participant), summaryh(lm(acc ~ percentOutC)), by = .(kmeansGroup)][term !="(Intercept)"]
```

Figure 4D
```{r figure relating percent out to classification separately for high vs. low lapsing adults}
ggplot(percents[!participant %in% c(outlier$participant) & group %in% ("Adults")], aes(percentOut*100, acc*100, col = as.factor(kmeansGroup)))+
  geom_point() +
  ggtheme+
  stat_smooth(data = percents[kmeansGroup == 1], formula = y ~ x, method = "lm") +
  stat_smooth(data = percents[kmeansGroup == 2], formula = y ~ x, method = "lm") +
  labs(x = "Lapse rates (%)", y = "Classification accuracy (%)") +
  scale_color_manual(values = c("blue4", "royalblue1")) +
  theme(legend.position = "None") +
  coord_cartesian(ylim = c(95, 100)) + 
  scale_y_continuous(breaks = c(94, 96, 98, 100), labels = c("94", "96", "98", "100"))
```

The following section is reported in analysis S3: Testing whether lapse rates mediate age group differences in classification accuracy
```{r eval = FALSE, include = FALSE, echo = FALSE}
#acme: mediated (ab) effect
#ade: direct effect (c')
#total effect: c
percents[!participant %in% c(outlier$participant), percentOutZ := scale(percentOut)]
percents[!participant %in% c(outlier$participant), mean_acc.allZ := scale(acc)]

model.y_all = lm(mean_acc.allZ ~ groupEC + percentOutZ , data = percents[!participant %in% c(outlier$participant)])
summary(model.y_all)

model.m_all = lm(percentOutZ ~ groupEC, data = percents[!participant %in% c(outlier$participant)])
summary(model.m_all)

model.mediate_attention_accuracy_all = mediate(model.m_all, model.y_all, sims = 5000, treat = 'groupEC', mediator = 'percentOutZ', covariates = 'mean_rt', boot = TRUE)
summary(model.mediate_attention_accuracy_all)
```

The following chunks of code fit models reported under the heading "Within-participant fluctuations in attention shape categorization accuracy similarly in children and adults" in the paper. These models test whether fluctuations in attention covary with classification accuracy. 
```{r eval = FALSE, include = FALSE, echo = FALSE}
# fit model relating fluctuations in attention to classification accuracy:
continuous_attention_accuracy_age_interaction.m = glmer(acc ~ pre_smoothed_deviance_scaled * trialTypeEC * groupEC + (pre_smoothed_deviance_scaled * trialTypeEC||participant), control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)), family  = "binomial", data = encoding); summaryh(continuous_attention_accuracy_age_interaction.m)
```

Figure 3d: Higher RT deviance predicts worse categorization:
```{r }
# for visualization purposes, separate the moving time course of RT deviance into a binary higher and lower deviance trials based on the median:
encoding[, median_pre_smoothed_deviance := median(pre_smoothed_deviance, na.rm=T), by = participant]
encoding[, pre_smoothed_deviance_cat := ifelse(pre_smoothed_deviance > median_pre_smoothed_deviance, "high", "low")]
encoding[pre_smoothed_deviance == median_pre_smoothed_deviance, pre_smoothed_deviance_cat := "low"]

#plot the data:
toPlot <- encoding[!is.na(pre_smoothed_deviance_cat), .(mean_acc = mean(acc, na.rm=T)), by = .(participant, group, pre_smoothed_deviance_cat)]
toPlot.w <- seWithin(data = toPlot, measurevar = "mean_acc", betweenvars = "group", withinvars = "pre_smoothed_deviance_cat", idvar = "participant")
ggplot(toPlot.w, aes(group, mean_acc*100, col = pre_smoothed_deviance_cat)) +
  #stat_summary()+
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_errorbar(aes(ymin = mean_acc*100 - se*100, ymax = mean_acc*100 + se*100), position = position_dodge(width = 0.5), width = 0, size = 1) + 
  ggtheme  + 
#scale_color_viridis(discrete = TRUE, option = "G", labels = c("Error", "Correct"))  +
  scale_color_manual(breaks = c("low", "high"), labels = c("low deviance", "high deviance"), values = c("orchid4", "seagreen4")) +
  scale_y_continuous(breaks = c(94, 96, 98, 100), labels = c("94", "96", "98", "100"))+
  coord_cartesian(ylim = c(94, 100)) +
  labs(y = "Accuracy (%)", x = " ", col = " ") 
```

The following code chunks are reported under the heading "Equivalent memory performance in children and adults"
```{r }
# read recognition file:
recognition <- fread(file = '../data/recognition.csv')

# compute high confidence d' for each participant:
Dprime.hc <- recognition[, dprime(hits = memoryHits_dprime, false_alarms = false_alarm_dprime, na.rm = T), by = .(participant, group, groupEC, groupA, groupC)]

# compute high confidence d' for each participant and trial type:
dprime.hc_trialtype <- recognition[, dprime(hits = memoryHits_dprime, false_alarms = false_alarm_dprime, na.rm = T), by = .(participant, group, groupEC, groupA, groupC, trialTypeI, trialTypeF, trialTypeEC)]

# scale d':
dprime.hc_trialtype[, dprimeC := scale(dprime, scale = F)]
```

# fit model relating age group to d':
```{r dprime by trial type}
dprime.hc_trialtype[!participant %in% c(outlier$participant), summaryh(lmer(dprime ~ groupEC * trialTypeEC + (1|participant)))][term != '(Intercept)']
dprime.hc_trialtype[!participant %in% c(outlier$participant), summaryh(lm(dprime ~ trialTypeEC + (1|participant))), by = group][term != '(Intercept)']
```

The following code reports analyses reported in the section "High lapse rates impair memory across children and adults who lapse often:

```{r join dprime to time out data}
# join dprime and percent out of the zone dataframe, lapse length and frequency
percents_dprime.hc = left_join(Dprime.hc, percents[, .N, by = .(participant, percentOut, percentOutC, groupEC)][, .(participant, percentOut, percentOutC)])

percents_dprime.hc_trialtype = left_join(dprime.hc_trialtype, percents[, .N, by = .(participant, percentOut, percentOutC, groupEC)][, .(participant, percentOut, percentOutC)])
```

Run correlations
```{r overall - collapsed across trial types}
# correlations the percent of time out of the zone with d'
percents_dprime.hc[!participant %in% c(outlier$participant), summaryh(lm(dprime ~ percentOutC*groupEC))][!term == "(Intercept)"]
percents_dprime.hc[!participant %in% c(outlier$participant), summaryh(lm(dprime ~ percentOutC)), by = group][!term == "(Intercept)"]
```

# Figure 4a: Attentional lapse rate and memory relationships by group

```{r figure attentional lapse rate and memory relationships by group}
ggplot(percents_dprime.hc[!participant %in% c(outlier$participant)], aes(percentOut *100, dprime, col = group))+
  geom_point() +
  ggtheme+
  stat_smooth(formula = y ~ x, method = "lm") +
  labs(x = "Time out of the zone (%)", y = "D prime") +
  #scale_x_continuous(breaks = c(0, .2, .4, .6, 0.8), labels = c(0, 20, 40, 60, 80)) +
  scale_y_continuous(breaks = c(1, 2, 3), labels = c(1, 2, 3)) +
  #coord_cartesian(ylim = c(1, 3), xlim = c(0, .8)) +
  scale_color_manual(breaks = c("Adults", "Children"), values = c("royalblue3", "hotpink3")) +
  theme(legend.position = "None")
```

The following section is reported under the heading "Fluctuations in sustained attention shape memory formation across time more in chidlren than adults:
```{r join memory and encoding data}
encoding = left_join(encoding, recognition[classifyMemory == 'old', .(participant, stimulus, memory_hits)])
```

```{r fit models}
# rt preceding deviance predicts memory hits
fluctuations_hits.m <- encoding[, summaryh(glmer(memory_hits ~ pre_smoothed_deviance_scaled  * classifyEncEC * groupEC +  (pre_smoothed_deviance_scaled * classifyEncEC ||participant), control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)), family = 'binomial'))][term != '(Intercept)']
print(fluctuations_hits.m)
```
